\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{parskip}

\setstretch{1.05}
\setlength{\parindent}{0pt}

\begin{document}

\begin{minipage}[b]{2.8in}
  Cover Letter \\
  {\bf GENETICS}
\end{minipage}
\hfill
\begin{minipage}[b]{2.8in}
  Hanbin Lee and Jonathan Terhorst \\
  \today
\end{minipage}

\vskip 2em

\noindent
To the Editor(s),

\vskip 1em

We are pleased to submit our manuscript,
``Parameterizing the genetic architecture under stabilizing selection.''
for consideration in the GENETICS special issue
``The Fitness Landscape: data, theory, and applications.''

This paper develops a theoretical and statistical framework linking
fitness-landscape models to mixed-model methodology in statistical genetics.
Starting from stabilizing selection under Fisher's geometric model,
we derive a frequency-dependent genetic architecture in which
effect-size variance emerges as a function of interpretable evolutionary quantities,
including mutational variance, selection intensity, and coupling between
the focal trait and the selected background trait.
The resulting model yields identifiable variance components that can be estimated by
restricted maximum likelihood and then used for downstream prediction by BLUP.

We believe the manuscript is well aligned with the aims of the special issue.
Although our application is motivated by human complex-trait genetics,
the central contribution is methodological and conceptual:
it treats the fitness landscape not merely as metaphor,
but as a quantitative object from which one can derive empirically testable structure
for genetic architecture and statistical inference.
In this sense, the manuscript speaks directly to ongoing efforts
to connect theoretical models of fitness landscapes with modern computational
and inferential practice.

We also expect the paper to be of interest to readers concerned with
how evolutionary constraints shape the distribution of genetic effects,
and with how such constraints can be incorporated into models used
for heritability estimation and prediction.
In forward simulations, the proposed framework recovers the focal-trait variance well
and generally improves predictive performance relative to conventional
$\alpha$-model baselines, while remaining grounded in an explicit population-genetic model.

This manuscript is original, is not under consideration elsewhere,
and has been approved by all authors.

\vskip 2em

\noindent
Sincerely,

\vskip 2em

Hanbin Lee, on behalf of the authors

\end{document}
